---
title: "STRING"
author: "Celine Hernandez"
date: "6 Jun 2016"
output: 
  html_document: 
    toc: yes
---

http://www.bioconductor.org/packages/3.3/bioc/vignettes/STRINGdb/inst/doc/STRINGdb.pdf

# Introduction

The STRING database is available at http://string-db.org

Description from the STRING website:
> STRING is a database of known and predicted protein-protein interactions. The interactions include direct (physical) and indirect (functional) associations; they stem from computational prediction, from knowledge transfer between organisms, and from interactions aggregated from other (primary) databases.

Remote access to the database can be achieved by either direct request to the API or by using the STRINGdb R library.


# STRING API

## Access using R

When not using the STRINGdb R library, it's necessary to use a generic access to the RESTful API through the RCurl library. 

RCurl: https://cran.r-project.org/web/packages/RCurl/index.html  

```{r install_library_rcurl, eval=FALSE}
# Not run

# Generic library used to 
install.packages("RCurl")
```

Once installed, we can load the library.

```{r load_rcurl}
library('RCurl')
```

To use RCurl more easily, we can define a generic function requesting a provided URL given a list() of parameters and returning the service's response.

```{r rcurl_request}
rcurl_request <- function(service_url, access, format, request, parameters) {
    
    # Collapse all parameters into one string
    all_parameters <- paste(
        sapply(names(parameters), 
               FUN=function(param_name, parameters) {
                   paste(param_name, paste(parameters[[param_name]], collapse='\n'), collapse='', sep='=')
               }, 
               parameters),
        collapse="&")
    
    # Paste base URL and parameters
    requested_url <- paste(paste(service_url, access, format, request, sep = "/"), all_parameters, sep="?")
    
    # Encode URL (in case there would be any space character for instance)
    requested_url <- URLencode(requested_url)
    
    # Start request to service
    response <- getURL(requested_url, .opts = list(ssl.verifypeer = FALSE))
    
    return(response)
}
```

## General description of the API

STRING API requests are based on a global synthax for the URL.

> http://[database]/[access]/[format]/[request]?[parameter]=[value]

See this image for the full detail of all combinations : http://string-db.org/help/images/api_format.png

### [database]

The STRING API provides two entry points (base URLs) to STRING as well as one access to another sister database called STITCH (chemical-protein interactions).

Database | Description
--- | ---
http://string-db.org | Main entry point to STRING
http://string.embl.de | Alternative entry point to STRING
http://stitch.embl.de | Entry point to STITCH

In this file, we will use the STRING main entry point.

```{r base_url}
STRING_service_url <- "http://string-db.org"
```


### [access]

Two types of access are provided, either "api" for the API, or "services" for other kind of services. (Which ones?)

In this file, we will focus on the API.

```{r access}
STRING_access <- "api"
```


### [format]

STRING provides either simple "table-like" tab-separated formats (TSV, TSV without headers or PSI-MI-TAB) as well as more complex formats like JSON or XML (PSI-MI format).

Format | Description
--- | ---
json | JSON format either as a list of hashes/dictionaries, or as a plain list (if there is only one value to be returned per record)
tsv | Tab separated values, with a header line
tsv-no-header | Tab separated values, without header line
psi-mi | The interaction network in PSI-MI 2.5 XML format
psi-mi-tab | Tab-delimited form of PSI-MI (similar to tsv, modeled after the IntAct specification. (Easier to parse, but contains less information than the XML format.)


To read the returned information, base R functions can be used directly. On the other hand, JSON data can be read using the RJSONIO R package, while the XML package is used for XML data.

RJSONIO: https://cran.r-project.org/web/packages/RJSONIO/index.html  
XML: https://cran.r-project.org/web/packages/XML/index.html  


```{r install_libraries, eval=FALSE}
# Not run

# If we want to access the information in JSON format
install.packages("RJSONIO")
# If we want to access the information in XML format (PSI-MI)
install.packages("XML")
```

```{r load_libraries}
library(RJSONIO)
library(XML)
```


### [request] and associated [parameter]=[value]

The [request] element of the URL is the type of request that needs to be performed on STRING data. For a given type of request, STRING provides two request names depending on if it's a request on one "identifier" (<requestname>) or on a list of "identifiers" (<requestname>List).

#### resolve / resolveList (ID matching)

Goal: match requested identifier(s) to STRING species. Be extremely careful as it matches not only on preferred names but also on the full annotation of each STRING entry.


Format: One of "json", "tsv" or "tsv-no-header". The same content is returned whatever the chosen format.


Allowed parameters | Description
--- | ---
identifier/identifiers | One or multiple identifiers (symbol or synonym or UniProt or ?)
format | One of "full" or "only-ids".
species | Either "auto-detect" or a taxonomic ID.


* Example of a request to "resolve" with one identifier (JSON).
Be aware that one should use "identifier" to provide one ID.

```{r resolve_ex1}
STRING_format <- "json"
STRING_request <- "resolve"
parameters <- list(identifier="Lat", #9606.ENSP00000378845
                   format="full",
                   species="9606")

# Perform request
result <- rcurl_request(STRING_service_url, STRING_access, STRING_format, STRING_request, parameters)

# Parse result
json_result <- fromJSON(result)
json_result <- data.frame(sapply(json_result, simplify = TRUE,
                      FUN=function(x) {
                          return(x[c("queryIndex", "stringId", "preferredName", "ncbiTaxonId", "taxonName", "annotation")])
                    }))

# Display result
print(json_result[1:5,])
```

* Example of a request to "resolveList" with a list of multiple identifier (TSV).  
Be aware that one should use "identifiers" and not "identifier" to provide the list of IDs.

```{r resolve_ex2}
STRING_format <- "tsv"
STRING_request <- "resolveList"
parameters <- list(identifiers=c("Lat", "Lck", "P07766"), # 9606.ENSP00000378845, 9606.ENSP00000337825
                   format="full",
                   species="9606")

# Perform request
result <- rcurl_request(STRING_service_url, STRING_access, STRING_format, STRING_request, parameters)

# Parse result
tsv_result <- read.table(text = result, header = TRUE, quote = "", stringsAsFactors = FALSE, row.names = NULL, sep = "\t")


# Display result
print(tsv_result[, 1:5])
```



#### abstracts / abstractsList

Goal: list abstracts matching requested identifier(s).


Format: One of "json", "tsv" or "tsv-no-header". The same content is returned whatever the chosen format.


Allowed parameters | Description
--- | ---
identifier/identifiers | One or multiple identifiers (symbol or synonym?)
format | One of "colon" or "pmids".
limit | Integer. Maximun number of returned abstracts.

* Example of a request to "abstracts" with one identifier (JSON).
Be aware that one should use "identifier" to provide one ID.

```{r abstracts_ex1}
STRING_format <- "json"
STRING_request <- "abstracts"
parameters <- list(identifier="9606.ENSP00000378845", #Lat
                   format="colon", # "pmids" returns nothing?? Other format: remove completely parameter "format"
                   limit="5")

# Perform request
result <- rcurl_request(STRING_service_url, STRING_access, STRING_format, STRING_request, parameters)

# Parse result
json_result <- fromJSON(result)

# Display result
print(json_result)
```

* Example of a request to "abstractsList" with a list of multiple identifier (TSV).  
Be aware that one should use "identifiers" and not "identifier" to provide the list of IDs.

```{r abstracts_ex2}
STRING_format <- "tsv"
STRING_request <- "abstractsList"
parameters <- list(identifiers=c("Lat", "Lck", "P07766"),
                   # "format" can be completely omitted
                   limit="5")

# Perform request
result <- rcurl_request(STRING_service_url, STRING_access, STRING_format, STRING_request, parameters)

# Parse result
tsv_result <- read.table(text = result, header = TRUE, quote = "", stringsAsFactors = FALSE, row.names = NULL, sep = "\t")


# Display result
print(tsv_result)
```



# Session info

```{r}
sessionInfo()
```

